Constructive Preference Elicitation by Setwise Max-margin Learning

### 2016-01-28 paper/ijcai16.tex
paper: update all figures with the latest axis labels.
paper: remove commented block.
paper: add machine desc.
paper: fix minor wart, remove todo.

\stefano{add early stopping in algorithm and description and experiments.}

@@

In Figure~\ref{fig:comparison} we report solution quality and timing
values for increasing number of collected user responses, for the
different competitors on each of the four different utility vector
distributions. Solution quality is measured in terms of utility loss:
%
$$ \max_{\vx\in\calX_\text{feasible}} \left( u(\vx) - u(\vx^*) \right) $$
%
%<<<<<<< HEAD
%varying the number of collected user responses. Here $u(\cdot)$ is the
%true unknown user utility, and $\vx^*$ is the solution recommended to
%the user after the elicitation phase (see Algorithm~\ref{alg:setmargin}).
%
%The thick line is the median
%utility loss taken over all 20 users, while the shaded area represents the
%standard deviation.
%
%\stefano{plot average utility loss at the increase of dataset size}
%
%\stefano{plot average runtime at the increase of dataset size}
%
%\stefano{comment on usefulness of sparse weights}
%
%\paolo{Maybe a plot of an experiment with increased noise in user reponse? (for example, $\beta=0.1$ instead of $1$.}
%
%\paragraph{Cognitive effort.} Contrarily to most other preference elicitation
%methods, which generate a single candidate option per iteration, {\sc
%SetMargin} constructs a whole {\em set} of options. We thus can not leverage
%standard query selection strategies \stefano{which ones?}. 
%
%The simplest solution is to query the user about all $k \choose 2$ pairs of
%configurations in $\{\vx^i\}$. Since $k$ is a small constant, the resulting
%number of queries does not grow large in practice.  An appealing alternative
%involves using a modified sorting algorithm, such as merge-sort, to order the
%set $\{\vx^i\}$ by asking only $O(n\log n)$ queries to the user. However, this
%approach is very susceptible to noise: a single inconsistent user response may
%affect all successive comparisons, leading to a cascading effect. While more
%robust strategies may be conceived, empirically we found that the simpler
%quadratic scheme tends to work reasonably well.
%
%\andrea{rewrite as footnote}
%=======
where $u(\cdot)$ is the true unknown user utility, and $\vx^*$ is the
solution recommended to the user after the elicitation phase (see
Algorithm~\ref{alg:setmargin}). Computational cost is measured in
terms of cumulative time. Given that Guo and Viappiani (both QI and
EUS) are single-threaded, we disabled multi-threading when running our
algorithm in these comparisons. All experiments were run on a XXXXXX\andrea{@stefano}.
In all figures, thick lines indicate median values over the different
users, shaded areas indicate standard deviations. Our {\sc SetMargin}
algorithm for $k=2$ is reported in orange, Guo in blue, Viappiani QI
in green and Viappiani EUS in gray. For all algorithms, one iteration
corresponds to a single pairwise query to the user.  Results for $r=3$
and $r=4$ are reported.  For dense weight vector distributions (first
two rows), our approach achieves results which are indistinguishable
from the competitors in a fraction of their time.  Indeed, all
Bayesian approaches become quickly intractable for growing values of $r$,
%cannot scale over $r=4$ (for $r=5$, Guo did not
%finish a single user after 24 hours for, Viappiani QI takes 20 minutes
%per user), \andrea{20 minuti mi paiono pochi, sei sicuro?}  
while our algorithm can easily scale to much larger datasets, as will
be shown later on. For sparse weight vector distributions (last two
rows) our approach, in addition to being substantially faster on each
iteration, requires less queries in order to reach optimal
solutions. This is an expected result as the sparsification norm in
our formulation ($\| \vw \|_1$) is enforcing sparsity in the weight
vectors, while none of the other approaches are explicitly designed to
favour sparse functions.

@@

In order to study the effect of increasing the number of weight
vectors in our formulation, we also ran {\sc SetMargin} varying the
parameter $k$. Figure~\ref{fig:selfcomparison} reports utility loss
results on $r=4$ and $r=5$ datasets for the uniform and sparse normal
distributions\footnote{We picked the most difficult dense distribution
  (uniform) and the simplest sparse distribution (normal) to provide
  the broadest possible picture, as space constraints prevent us from
  including all results. Results for the other distributions are
  however qualitative similar to the ones reported.}, for $k=2$
(orange), $k=3$ (blue) and k=4 (green). The first and third columns
report results in terms of number of iterations. It can be seen that
increasing the number of weight vectors tends to favour earlier
convergence, especially for the more complex dataset ($r=5$). However,
as in each iteration the user is asked to compare $k$ items, different
values of $k$ imply a different cognitive effort for the user. The
second and fourth columns report results in terms of number of
queries, where we count all $k \choose 2$ pairs of queries when
comparing $k$ items. In this case, $k=2$ seems to be the best
option. The cognitive cost for the user will likely lay in between
these two extremes, but formalizing this concept in an efficient query
ordering strategy needs to face the effect of noise.  A modified
sorting algorithm asking only $O(k\log k)$ queries to the user
resulted in a performance worsening, likely because of a cascading
effect of inconsistence feedback. \andrea{@stefano: l'ho rifrasato un
  po', vediamo se ci piace}

>>>>

@@

In Figure~\ref{fig:comparison} we report solution quality and timing
values for increasing number of collected user responses, for the
different competitors on each of the four different utility vector
distributions. Solution quality is measured in terms of utility loss:
%
$$ \max_{\vx\in\calX_\text{feasible}} \left( u(\vx) - u(\vx^*) \right) $$
%
where $u(\cdot)$ is the true unknown user utility, and $\vx^*$ is the
solution recommended to the user after the elicitation phase (see
Algorithm~\ref{alg:setmargin}). Computational cost is measured in
terms of cumulative time. Given that Guo and Viappiani (both QI and
EUS) are single-threaded, we disabled multi-threading when running our
algorithm in these comparisons. All experiments were run on a 2.8 GHz Intel
Xeon CPU with 8 cores and 32 GiB of RAM.
In all figures, thick lines indicate median values over the different
users, shaded areas indicate standard deviations. Our {\sc SetMargin}
algorithm for $k=2$ is reported in orange, Guo in blue, Viappiani QI
in green and Viappiani EUS in gray. For all algorithms, one iteration
corresponds to a single pairwise query to the user.  Results for $r=3$
and $r=4$ are reported.  For dense weight vector distributions (first
two rows), our approach achieves results which are indistinguishable
from the competitors in a fraction of their time.  Indeed, all
Bayesian approaches become quickly intractable for growing values of $r$,
%cannot scale over $r=4$ (for $r=5$, Guo did not
%finish a single user after 24 hours for, Viappiani QI takes 20 minutes
%per user), \andrea{20 minuti mi paiono pochi, sei sicuro?}  
while our algorithm can easily scale to much larger datasets, as will
be shown later on. For sparse weight vector distributions (last two
rows) our approach, in addition to being substantially faster on each
iteration, requires less queries in order to reach optimal
solutions. This is an expected result as the sparsification norm in
our formulation ($\| \vw \|_1$) is enforcing sparsity in the weight
vectors, while none of the other approaches are explicitly designed to
favour sparse functions.

@@

In order to study the effect of increasing the number of weight
vectors in our formulation, we also ran {\sc SetMargin} varying the
parameter $k$. Figure~\ref{fig:selfcomparison} reports utility loss
results on $r=4$ and $r=5$ datasets for the uniform and sparse normal
distributions\footnote{We picked the most difficult dense distribution
  (uniform) and the simplest sparse distribution (normal) to provide
  the broadest possible picture, as space constraints prevent us from
  including all results. Results for the other distributions are
  however qualitative similar to the ones reported.}, for $k=2$
(orange), $k=3$ (blue) and $k=4$ (green). The first and third columns
report results in terms of number of iterations. It can be seen that
increasing the number of weight vectors tends to favour earlier
convergence, especially for the more complex dataset ($r=5$). However,
as in each iteration the user is asked to compare $k$ items, different
values of $k$ imply a different cognitive effort for the user. The
second and fourth columns report results in terms of number of
queries, where we count all $k \choose 2$ pairs of queries when
comparing $k$ items. In this case, $k=2$ seems to be the best
option. The cognitive cost for the user will likely lay in between
these two extremes, but formalizing this concept in an efficient query
ordering strategy needs to face the effect of noise.  A modified
sorting algorithm asking only $O(k\log k)$ queries to the user
resulted in a performance worsening, likely because of a cascading
effect of inconsistence feedback.
